---
title: "Volcano Plots"
author: "Rob Lampe"
date: "1/29/2022"
---

Example code for making volcano plot

```{r}
library(ggplot2)
library(scales)
```

Read in results from DESeq2.
The columns of interest from DESeq2 results are log2FoldChange and padj,
but these can be substituted from results from other calculations

```{r}
res <- read.csv('deseq2_results.csv')
```

static vars

```{r}
max_log_p <- 5
alpha <- 0.05
```

Remove points where padj = NA and log10 transform p-values
See DESeq2 documentation for explanation: 
https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#pvaluesNA

```{r}
res<- res[!is.na(res$padj),]
res$log_p <- -log10(res$padj)
```

Set max -log10(P-value) to 5

```{r}
res[res$log_p > max_log_p,]$log_p <- max_log_p 
```

Function for reversed log10 scale
Source: https://stackoverflow.com/a/11054781

```{r}
reverselog_trans <- function(base = exp(1)) {
  trans <- function(x) -log(x, base)
  inv <- function(x) base^(-x)
  trans_new(paste0("reverselog-", format(base)), trans, inv, 
            log_breaks(base = base), 
            domain = c(1e-100, Inf))
}
```

Plot

```{r}
ggplot(res, aes(x=log2FoldChange, y=log_p)) + 
  geom_point(data = res[res$padj > alpha,], aes(x=log2FoldChange, y=log_p), color="grey", size = 2.5) + # points with padj > alpha are grey 
  geom_point(data = res[res$padj <= alpha & res$log_p < max_log_p,], aes(x=log2FoldChange, y=log_p), size = 2.5) + # point with padj < alpha and log10(padj) < max_log_p are black circles
  geom_point(data = res[res$log_p == max_log_p,], aes(x=log2FoldChange, y=log_p), size = 2.5, shape = 17) + # points with padj < alpha and log10(padj) = max_log_p are black triangles
  geom_vline(xintercept = 0, linetype = 5) +
  geom_hline(yintercept = -log10(alpha), linetype = 5) + 
  ylim(0, 5) +
  scale_x_continuous(breaks = seq(-10, 10, by = 2), limits=c(-12, 12)) +
  theme_bw() +
  xlab(bquote(~log[2]~"fold change")) + ylab(expression(paste("-log(adjusted ", italic("p"), "-value)", sep=""))) +
  theme(panel.border = element_rect(colour = "black"), 
        axis.line = element_line(colour = "black"), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title=element_text(size=18), 
        axis.text = element_text(size=16, colour="black"),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

uncomment to save as pdf
```{r}
#ggsave("volcano_plot.pdf", width = 6, height = 8)
```
